Whole-field total density measurements by Digital Image Correlation \thanks{Grants or other notes about the article that should go on the front page should be placed here. }

### 2020-02-12 Article/experimentsInFluids.tex
Journal Format

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{subcaption}
\usepackage{standalone}
\usepackage{xcolor}
\usepackage{enumitem}
\setlist{nosep}

\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.pathreplacing}

\usepackage{footnote}
\usepackage{csquotes}
\usepackage[USenglish]{babel}
\usepackage[backend=biber]{biblatex}

\addbibresource{articleBOS.bib}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
\DeclareMathOperator{\sign}{sign}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Insert your title here%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\institute{Alexander M. van Oers \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Leo R.M. Maas \at
              second address
}

@@

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

In a typical BOS or SS application, we image perturbations, or waves, traveling through a fluid. Two or more images are compared using Digital Image Correlation (DIC) [refs]. This comparison yields the displacements between the images. These displacements are related to the changes in the gradients of the index of refraction using a ray tracing model. These gradients of the index of refraction are related to the gradients of the density by experimentally determined data [refs], by the Gladstone-Dale relation for gases [refs] or by the Lorentz-Lorenz equation for liquids [refs]. The density field is sometimes obtained from the gradients of the density perturbations by solving a Poisson equation [refs].

In this paper we present a modification to the BOS or SS technique to directly measure the density field. We take images of static situations:  (1) before filling a tank in the experimental setup with (stratified) fluids, (2) after filling the tank with a calibration fluid and (3) after filling the tank with (stratified) fluids. Then we take images of the dynamic situations: we image, as in BOS or SS, the perturbations, or waves, traveling through a fluid. We perform DIC to obtain displacements between the static images. Using a ray tracing model we relate these displacements to the index of refraction. These static images give us the background index of refraction. We also perform DIC to obtain the displacements between the dynamic images and a static image.

The challenge of this new method was to have enough accuracy in our measurements and models to ensure that our results were useful. Determining the value of the density is more sensitive to noise than determining the density gradients of perturbations in a BOS or SS application. We also wanted to preserve the experimental simplicity of the BOS or SS method. In Section \ref{sec:simmod} we perform a naive application of the BOS or SS technique to try to measure the index of refraction of homogeneous water. We analyze a simple ray tracing model to determine why our naive approach failed. We show that light rays, traveling from the background through the fluid to the camera, should not travel in nearly straight lines (as in BOS or SS) but should have large deflections. We should either rotate the fluid tank or rotate (and move) our camera. We should not be placing our camera right in front of the fluid tank.

To obtain the required accuracy in determining the density field we optimized all the steps in our method. In Section \ref{sec:formod} we derive our ray tracing model relating displacements to the index of refraction. We call this our forward model: given an index of refraction field and experimental parameters, we can compute the displacement field of our experimental setup. This ray tracing model is a 3D model that allows the camera to be placed under an angle with respect to the fluid tank. In Section \ref{sec:cal} we describe our calibration procedure. This calibration step allows us to determine very accurately certain parameters in our forward model.  In Section \ref{sec:DIC} we discuss DIC. We provide references to papers that discuss techniques that allowed us to get the most of our DIC procedure. We also provide information about our settings for a typical DIC calculation. In Section \ref{sec:exp} we briefly discuss the experimental setup. We also discuss best practices to get the most out of our experiments. In Section \ref{sec:invmod} we discuss how to solve our inverse model: how to obtain the index of refraction from the experimentally obtained displacements and our forward model.

@@

\begin{figure}[hpbt]
	\includestandalone{schematicsetupangle}
	\caption{A schematic view of a path of a light ray (not to scale). The light ray in the reference state is indicated by a solid line, the light ray after filling the tank with water by a dashed line. The numbers indicate the planes through which the light ray propagates: 0, the ccd sensor inside the camera; 1, the camera pinhole; 2, the start of the first glass plate; 3, the end of the first glass plate; 4, the start of the second glass plate; 5, the end of the second glass plate; 6, the screen from which the light rays originate.  $n_0$, $n_1$ and $n_2$ are the index of refraction of, respectively, air, material of the tank and the fluid. The lengths (in the $z$-direction) are: $L_c$, the distance from camera to the first glass plate; $L_g$, the width of the glass plates; $L_t$ the width of the tank; $L_s$, the distance from the second glass plate to the screen.}	
\label{fig:schviepalira}	
\end{figure}

Assuming the refractive index in each section of the set up is constant, we can write for the light path,
\begin{equation}
	\label{eq:simpleline}
x(z) = x_i + z \tan \theta_i, \qquad i = 1, ... , 5,
\end{equation}
where $x_i$ is the $x$-position of the light ray at the start of each section and $\theta_i$ is the angle between the direction the ray is propagating and the $z$-axis. When encountering a discontinuous change in refractive index (when passing between the sections), we invoke Snell's laws
\begin{equation}
	\label{eq:snellslaw}
n_i \sin \theta_i = \mbox{ constant},
\end{equation}
where $n_i$ is the refractive index for the $i^{th}$ section. The position $x_6$ is
\begin{align}
\label{eq:simplex6}
x_6 (\theta_x, n_2) & =  (L_c+L_s) \tan \theta_x + 2 L_g \tan \arcsin \frac{n_0}{n_1} \sin \theta_x + L_t \tan \arcsin \frac{n_0}{n_2} \sin \theta_x \nonumber \\
& =  (L_c+L_s) \tan \theta_x + \frac{2 L_g n_0 \sin \theta_x}{\sqrt{n_1^2 - n_0^2 \sin^2 \theta_x}} + \frac{L_t n_0 \sin \theta_x}{\sqrt{n_2^2 - n_0^2 \sin^2 \theta_x}}.
\end{align}
The displacement $\Delta x$ is the difference in position $x_6$ between the constant reference state, $n_2 = n_0$, and the unknown state $n_2=n$, 
\begin{align}
\label{eq:dexcon2}
\Delta x = x_6 (n_2 = n_0) - x_6 (n_2 = n) & = L_t \left[ \tan \theta_x - \tan \arcsin \frac{n_0}{n} \sin \theta_x\right] \nonumber \\
& = L_t \left[ \tan \theta_x - \frac{n_0 \sin \theta_x}{\sqrt{n^2 - n_0^2 \sin^2 \theta_x}} \right] .
\end{align}
Inverting this relation we obtain the unknown refractive index $n$ as a function of the displacement
\begin{equation}
\label{eq:invdexcon2}
n = \frac{n_0 \sin \theta_x}{\sin \arctan \left( \tan \theta_x - \frac{\Delta x}{L_t} \right)} = \frac{n_0 \sin(\theta_x)}{\tan(\theta_x)-\frac{\Delta x}{L_t}} \sqrt{1+\left(\tan(\theta_x)-\frac{\Delta x}{L_t}\right)^2}.
\end{equation}
The angle $\theta_x$, the displacement $\Delta x$ and unknown index of refraction $n$ are functions of the coordinate $x$. % Equation (\ref{eq:invdexcon2}) holds for each point.

%The index of refraction of air $n_0$ is constant. The light ray hits the first glass plate with a position
%\begin{equation}
%\label{eq:tanlaw}
%x_2 = L_c \tan \theta_x.
%\end{equation}
%The index of refraction of the glass plate $n_1$ is constant. The angle under which the light ray propagates changes instantaneously, obeying Snell's law
%\begin{equation}
%\label{eq:snelllaw}
%n_0 \sin \theta_x = n_1 \sin \theta_T.
%\end{equation}
%For a constant index of refraction in the water tank $n_2$ the position $x_6$ is
%\begin{equation}
%x_6 (\theta_x, n_2) = (L_c+L_s) \tan \theta_x + 2 L_g \tan \arcsin \frac{n_0}{n_1} \sin \theta_x + L_t \tan \arcsin \frac{n_0}{n_2} \sin \theta_x
%\end{equation}
%The displacement $\Delta x$ is the difference in position $x_6$ for two different values of $n_2$. We take one known value, the %tank is filled with air, $n_0$, and one unknown value $n_2 = n$. 
%\begin{equation}
%\label{eq:dexcon2}
%\Delta x = x_6 (n_2 = n_0) - x_6 (n_2 = n) = L_t \left[ \tan \theta_x - \tan \arcsin \frac{n_0}{n} \sin \theta_x\right].
%\end{equation}
%We can invert this relation to obtain the unknown $n$,
%\begin{equation}
%\label{eq:nfromdxsimple}
%	n = \frac{n_0 \sin(\theta_x)}{\tan(\theta_x)-\frac{\Delta x}{L_t}} \sqrt{1+\left(\tan(\theta_x)-\frac{\Delta x}{L_t}\right)^2}.
%\end{equation}

Figure \ref{fig:simmod} shows the results when using (\ref{eq:invdexcon2}) to obtain the index of refraction $n$. We have two frontal images of the experimental setup; (1) Figure \ref{fig:air0simplefrontal}, without water, and (2) Figure \ref{fig:fresh0simplefrontal}, with water.  Applying a Digital Image Correlation (DIC) method we obtain (1) the correlation coefficient, a measure of the reliability of the method, (2) the horizontal displacements $\Delta x$, and (3) the vertical displacements $\Delta y$. Where coordinates where $\Delta x = 0$ and $\Delta y = 0$ are the location of the $z-axis$. We calculate the angles $\theta_x$ and $\theta_y$ with respect to this $z$-axis. Using (\ref{eq:invdexcon2}) we can compute the index of refraction $n$. Figure \ref{fig:nfromdxsimplefrontal} shows $n$ obtained from the horizontal displacements $\Delta x$.  Figure \ref{fig:nfromdysimplefrontal} shows $n$ obtained from the vertical displacements $\Delta y$. In Figure \ref{fig:nfrontal} we combine the horizontal and vertical displacements to obtain $n$.

\begin{figure}[htbp]
\begin{subfigure}{.45\linewidth}
		\centering \includegraphics[width = \textwidth,keepaspectratio]{Refsimplefrontal}
		\subcaption{Reference Image: Filled with air}\label{fig:air0simplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{deformedsimplefrontal}
	\subcaption{Deformed Image: Filled with water}\label{fig:fresh0simplefrontal}
\end{subfigure}\\
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth]{CCsimplefrontal}
	\subcaption{Correlation Coefficient from DIC}\label{fig:CCsimplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{DXsimplefrontal}
	\subcaption{$\Delta x$ from DIC}\label{fig:DXsimplefrontal}
\end{subfigure}\\
	\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{DYsimplefrontal}
	\subcaption{$\Delta y$ from DIC}\label{fig:DYsimplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{nfromdxsimplefrontal}
	\subcaption{$n$ from $\Delta x$}\label{fig:nfromdxsimplefrontal}
\end{subfigure}\\
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{nfromdysimplefrontal}
	\subcaption{$n$ from $\Delta y$}\label{fig:nfromdysimplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{nsimplefrontal}
	\subcaption{$n$ from $\sqrt{\Delta x^2+\Delta y^2}$}\label{fig:nfrontal}
\end{subfigure}
\caption{Simple Model Result. When the angle $\theta_x$ is near zero, large errors in $n$ appear. }
\label{fig:simmod}
\end{figure}

We see that when the displacements approach zero, we have very large errors in $n$. Figure \ref{fig:nfrontal} looks best, but is still unacceptable. Around the location of the $z-axis$, where the angles and displacements approach zero, the errors are very large. The entire figure should have a constant value of $n = 1.333$. This is not the case.

@@

To solve this, we want $\theta_x$ to not approach zero. Then, according to (\ref{eq:dexcon2}), $\Delta x$ does not approach zero and our calculation of $n$ is not plagued by a fraction that approaches $0/0$. To achieve this, we place our camera at an angle $\alpha$ with respect to the water tank. The angle $\theta_x$ is replaced by $\alpha+\phi_x$, where $\phi_x$ is the angle of the light ray relative to the principal axis of the camera, which is at an angle $\alpha$. For example, we position our camera at an angle of $\alpha=30^\circ$ relative to the tank. The camera has a viewing range of $[-4,4]$ degrees. Then $\phi_x$ varies from $-4^\circ$ to $4^\circ$.

@@

Since there is also an uncertainty in the angle, we want $n$ to vary with $\theta_x$ as little as possible. Figures \ref{fig:dndt} show this derivative. The magnitude of this derivative $\left|\partial n/\partial \theta_x\right|$ first decreases with increasing $\theta_x$, reaches a minimum, then increases again. The worst possible angles to measure at are $0^\circ$ and $90^\circ$. This minimum is reached at $31^\circ$. Any angle between $15^\circ$ and $60^\circ$ yields approximately the same sensitivity of $n$.

@@

%To reduce

Concluding, Figures \ref{fig:dndx} and \ref{fig:dndt} indicate that the worst angles to measure are close to zero. Angles between $15^\circ$ and $60^\circ$ result in small errors in $n$.

\clearpage 
\section{Forward Model}
\label{sec:formod}
In this section we build a forward model, relating an index of refraction field $n$, position on the ccd-sensor $\underline{x}$ and parameters $\underline{\alpha}$ to the coordinates $\underline{X}$ where the light rays hits the screen. We seek a model $\underline{X}(n, \underline{\alpha}, \underline{x})$. This is an extension of (\ref{eq:simplex6}).

In Subsection \ref{subsec:rayequ} we derive the equations governing 3-dimensional ray optics in an inhomogeneous medium. In Subsection \ref{subsec:RayPath} we use these equations to construct the full ray path from the camera through the tank to the screen.

@@

\subsection{Relating Refractive Index to Ray Path}
\label{subsec:RayPath}
To give our method flexibility and not making assumption on the relative configuration in our experimental setup, we model the light rays as traversing in a three-dimensional space and allow the different elements in the experimental setup to have any alignments.

@@

\subsubsection{Snell's Law}
To find the angle of incidence, $\theta_I$, between the incoming light ray and the plane, we compute
\begin{equation}
	\cos \theta_I = \underline{\hat{n}} \cdot \underline{I}.
\end{equation}
We apply Snell's law (\ref{eq:snellslaw}) to find the angle of refraction, $\theta_T$. The index of refraction of the medium through which the incoming ray travels is $n_I$, the index of refraction of the medium through wich the outgoing ray travels is $n_T$. To find the direction cosines of the outgoing ray, $\underline{T}$, we calculate 
\begin{align}
	S &= \sign \cos \theta_I, \\
	\sin \theta_T &= \frac{n_I}{n_T} \sqrt{1-\cos^2 \theta_I}, \\
	\cos \theta_T &= \sqrt{1-\left(\frac{n_I}{n_T}\right)^2(1-\cos^2 \theta_I)} \\
	\underline{T} &= \frac{n_I}{n_T} \underline{I} + \left(\frac{n_I}{n_T} \cos \theta_I - S \cos\theta_T\right)\underline{\hat{n}},	
\end{align}
where $S$, the sign of $\cos\theta_I$ ensures we stay in the right quadrant.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\RequirePackage{fix-cm}
%
\documentclass[draft]{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{geometry}
\usepackage{subcaption}

%\usepackage{xcolor}
\usepackage{enumitem}
\setlist{nosep}
\usepackage{footnote}
%%%
% Remove these when submitting
% We can turn the standalone tikz figures into regular figure files
%%%
\usepackage{standalone}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.pathreplacing}

\usepackage{csquotes}
\usepackage[USenglish]{babel}
\usepackage[backend=biber]{biblatex}

\addbibresource{articleBOS.bib}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
\DeclareMathOperator{\sign}{sign}
%
% Insert the name of "your journal" with
 \journalname{Experiments in Fluids}
%
\begin{document}

\title{Whole-field total density measurements by Digital Image Correlation
\thanks{Grants or other notes
about the article that should go on the front page should be
placed here. }
}

%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\institute{Alexander M. van Oers \at
             NIOZ Royal Netherlands Institute for Sea Research and Utrecht University, P.O. Box 59, 1790 AB Texel, The Netherlands \\
             \emph{Present address:} Netherlands Defence Academy, The Netherlands  %  if needed
           \and
           Leo R.M. Maas \at
              Imau, Inst. for Marine and Atmospheric research Utrecht, Utrecht University, Princetonplein 5, 3584 CC, Utrecht, The Netherlands
}

@@

\begin{abstract}
An optical method for the quantitative measurement of the total density field of two-dimensional stratified or homogeneous transparent fluids is presented. It is based on the Synthetic Schlieren method. Application of this method is illustrated by the determination of the static background density of a two-layer fluid and of a stratified fluid. Further aspects of the technique are illustrated by considering the dynamic total density field of a wave attractor in a stratified fluid.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

In a typical BOS or SS application, perturbations (waves) in a fluid contained in a tank are imaged. Two or more images are compared using Digital Image Correlation (DIC) \cite{sutton2009image}. This comparison yields the displacements between the images. These displacements are related to the changes in the gradients of the index of refraction using a ray tracing model. These gradients of the index of refraction are related to the gradients of the density by experimentally determined data (e.g. \cite{tan2015dependence}), by the Gladstone-Dale relation for gases \cite{born2013principles} or by the Lorentz-Lorenz equation for liquids \cite{lorentz1916theory}. The density field is sometimes obtained from the gradients of the density perturbations by solving a Poisson equation \cite{venkatakrishnan2004density, verso2015background}.

In this paper we present a modification to the BOS or SS technique to directly measure the density field. We take images of static situations:  (1) before filling a tank in the experimental setup with (stratified) fluids, (2) after filling the tank with a calibration fluid and (3) after filling the tank with (stratified) fluids. Then we take images of the dynamic situations: we image, as in BOS or SS, the perturbations, or waves, traveling through a fluid. We perform DIC to obtain displacements between the static images. Using a ray tracing model we relate these displacements to the index of refraction. These static images give us the background index of refraction. We follow the same procedure for the dynamic images.

The challenge of this new method is to have enough accuracy in our measurements and models to ensure that our results are useful. Determining the value of the density is more sensitive to noise than determining the density gradients of perturbations in a BOS or SS application. We also want to preserve the experimental simplicity of the BOS or SS method. In Section \ref{sec:simmod} we perform a naive application of the BOS or SS technique to try to measure the index of refraction of homogeneous water. We analyze a simple ray tracing model to determine why our naive approach failed. We show that light rays, traveling from the background through the fluid to the camera, should not travel in nearly straight lines (as in BOS or SS) but should have large deflections. We should either rotate the fluid tank or rotate (and move) our camera. We should not be placing our camera right in front of the fluid tank.

To obtain the required accuracy in determining the density field we optimize all the steps in our method. In Section \ref{sec:formod} we derive our ray tracing model relating displacements to the index of refraction. We call this our forward model: given an index of refraction field and experimental parameters, we can compute the displacement field of our experimental setup. This ray tracing model is a 3D model that allows the camera to be placed under an angle with respect to the fluid tank. In Section \ref{sec:cal} we describe our calibration procedure. This calibration determines accurately certain parameters in our forward model.  In Section \ref{sec:DIC} we discuss DIC. We provide references to papers that discuss techniques that allow us to get the most of our DIC procedure. We also provide information about our settings for a typical DIC calculation. In Section \ref{sec:exp} we briefly discuss the experimental setup. We also discuss best practices to get the most out of our experiments. In Section \ref{sec:invmod} we discuss how to solve our inverse model: how to obtain the index of refraction from the experimentally obtained displacements and our forward model.

@@

\begin{figure}[hpbt]
	\includestandalone[width=\textwidth]{schematicsetupangle}
	\caption{A schematic view of a path of a light ray (not to scale). The light ray in the reference state is indicated by a solid line, the light ray after filling the tank with water by a dashed line. The numbers indicate the planes through which the light ray propagates: 0, the ccd sensor inside the camera; 1, the camera pinhole; 2, the start of the first glass plate; 3, the end of the first glass plate; 4, the start of the second glass plate; 5, the end of the second glass plate; 6, the screen from which the light rays originate.  $n_0$, $n_1$ and $n_2$ are the index of refraction of, respectively, air, material of the tank and the fluid. The lengths (in the $z$-direction) are: $L_c$, the distance from camera to the first glass plate; $L_g$, the width of the glass plates; $L_t$ the width of the tank; $L_s$, the distance from the second glass plate to the screen.}	
\label{fig:schviepalira}	
\end{figure}

Assuming the refractive index in each section of the set up is constant, we can write for the light path,
\begin{equation}
	\label{eq:simpleline}
x(z) = x_i + z \tan \theta_i, \qquad i = 1, ... , 5,
\end{equation}
where $x_i$ is the $x$-position of the light ray at the start of each section and $\theta_i$ is the angle between the direction the ray is propagating and the $z$-axis. When encountering a discontinuous change in refractive index (when passing between the sections), we invoke Snell's laws
\begin{equation}
	\label{eq:snellslaw}
n_i \sin \theta_i = \mbox{ constant},
\end{equation}
where $n_i$ is the refractive index for the $i^{th}$ section. The position $x_6$ is
%\begin{align}
%\label{eq:simplex6}
%x_6 (\theta_x, n_2) & =  (L_c+L_s) \tan \theta_x + 2 L_g \tan \arcsin \frac{n_0}{n_1} \sin \theta_x + L_t \tan \arcsin \frac{n_0}{n_2} \sin \theta_x \nonumber \\
%& =  (L_c+L_s) \tan \theta_x + \frac{2 L_g n_0 \sin \theta_x}{\sqrt{n_1^2 - n_0^2 \sin^2 \theta_x}} + \frac{L_t n_0 \sin \theta_x}{\sqrt{n_2^2 - n_0^2 \sin^2 \theta_x}}.
%\end{align}
\begin{equation}
\label{eq:simplex6}
x_6 (\theta_x, n_2) =  (L_c+L_s) \tan \theta_x + \frac{2 L_g n_0 \sin \theta_x}{\sqrt{n_1^2 - n_0^2 \sin^2 \theta_x}} + \frac{L_t n_0 \sin \theta_x}{\sqrt{n_2^2 - n_0^2 \sin^2 \theta_x}}.
\end{equation}
The displacement $\Delta x$ is the difference in position $x_6$ between the constant reference state, $n_2 = n_0$, and the unknown state $n_2=n$, 
%\begin{align}
%\label{eq:dexcon2}
%\Delta x = x_6 (n_2 = n_0) - x_6 (n_2 = n) & = L_t \left[ \tan \theta_x - \tan \arcsin \frac{n_0}{n} \sin \theta_x\right] \nonumber \\
%& = L_t \left[ \tan \theta_x - \frac{n_0 \sin \theta_x}{\sqrt{n^2 - n_0^2 \sin^2 \theta_x}} \right] .
%\end{align}
\begin{equation}
\label{eq:dexcon2}
\Delta x = x_6 (n_2 = n_0) - x_6 (n_2 = n) = L_t \left( \tan \theta_x - \frac{n_0 \sin \theta_x}{\sqrt{n^2 - n_0^2 \sin^2 \theta_x}} \right) .
\end{equation}
Inverting this relation we obtain the unknown refractive index $n$ as a function of the displacement
\begin{equation}
\label{eq:invdexcon2}
n =\frac{n_0 \sin(\theta_x)}{\tan(\theta_x)-\frac{\Delta x}{L_t}} \sqrt{1+\left(\tan(\theta_x)-\frac{\Delta x}{L_t}\right)^2}. % \frac{n_0 \sin \theta_x}{\sin \arctan \left( \tan \theta_x - \frac{\Delta x}{L_t} \right)} = 
\end{equation}
The angle $\theta_x$, the displacement $\Delta x$ and unknown index of refraction $n$ are functions of the coordinate $x$. % Equation (\ref{eq:invdexcon2}) holds for each point.

Figure \ref{fig:simmod} shows the results when using (\ref{eq:invdexcon2}) to obtain the index of refraction $n$. We have two frontal images of the experimental setup; (1) Figure \ref{fig:air0simplefrontal}, without water, and (2) Figure \ref{fig:fresh0simplefrontal}, with water.  Applying DIC we obtain; Figure \ref{fig:CCsimplefrontal} the correlation coefficient, a measure of the reliability of the method; Figure \ref{fig:DXsimplefrontal} the horizontal displacements $\Delta x$; and (3) Figure \ref{fig:DYsimplefrontal} the vertical displacements $\Delta y$. Where $\Delta x = \Delta y = 0$ is the location of the $z$-axis. We calculate the angles $\theta_x$ and $\theta_y$ with respect to this $z$-axis. Using (\ref{eq:invdexcon2}) we can compute the index of refraction $n$. Figure \ref{fig:nfromdxsimplefrontal} shows $n$ obtained from the horizontal displacements $\Delta x$.  Figure \ref{fig:nfromdysimplefrontal} shows $n$ obtained from the vertical displacements $\Delta y$. In Figure \ref{fig:nfrontal} we combine the horizontal and vertical displacements to obtain $n$.

\begin{figure}[htbp]
\begin{subfigure}{.45\linewidth}
		\centering \includegraphics[width = \textwidth,keepaspectratio]{Refsimplefrontal}
		\subcaption{Reference Image: Filled with air}\label{fig:air0simplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{deformedsimplefrontal}
	\subcaption{Deformed Image: Filled with water}\label{fig:fresh0simplefrontal}
\end{subfigure}\\
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth]{CCsimplefrontal}
	\subcaption{Correlation Coefficient from DIC}\label{fig:CCsimplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{DXsimplefrontal}
	\subcaption{$\Delta x$ from DIC}\label{fig:DXsimplefrontal}
\end{subfigure}\\
	\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{DYsimplefrontal}
	\subcaption{$\Delta y$ from DIC}\label{fig:DYsimplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{nfromdxsimplefrontal}
	\subcaption{$n$ from $\Delta x$}\label{fig:nfromdxsimplefrontal}
\end{subfigure}\\
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{nfromdysimplefrontal}
	\subcaption{$n$ from $\Delta y$}\label{fig:nfromdysimplefrontal}
\end{subfigure}%
\begin{subfigure}{.45\linewidth}
	\centering \includegraphics[width = \textwidth,keepaspectratio]{nsimplefrontal}
	\subcaption{$n$ from $\sqrt{\Delta x^2+\Delta y^2}$}\label{fig:nfrontal}
\end{subfigure}
\caption{Simple Model Result. Applying DIC to the Reference Image in \ref{fig:air0simplefrontal} and the Deformed Image \ref{fig:fresh0simplefrontal} yields the Correlation Coefficient in \ref{fig:CCsimplefrontal}, the horizontal displacements in \ref{fig:DXsimplefrontal} and the vertical displacements in \ref{fig:DYsimplefrontal}. Using (\ref{eq:invdexcon2}) the index of refraction $n$ in \ref{fig:nfromdxsimplefrontal}, \ref{fig:nfromdysimplefrontal} and \ref{fig:nfrontal} are obtained. When the displacements $\Delta x$ and $\Delta y$ approach zero, large errors in $n$ appear. }
\label{fig:simmod}
\end{figure}

We see that when the displacements approach zero, we have very large errors in $n$. Figure \ref{fig:nfrontal} looks best, but is still unacceptable. Around the location of the $z$-axis, where the angles and displacements approach zero, the errors are very large. The entire figure should have a constant value of $n = 1.333$. This is not the case.

@@

To solve this, we want $\theta_x$ to not approach zero. Then, according to (\ref{eq:dexcon2}), $\Delta x$ does not approach zero and our calculation of $n$ is not plagued by a fraction that approaches $0/0$. To achieve this, we place our camera at an angle $\alpha$ with respect to the water tank. The angle $\theta_x$ is replaced by $\alpha+\phi_x$, where $\phi_x$ is the angle of the light ray relative to the principal axis of the camera, which is at an angle $\alpha$. %For example, we position our camera at an angle of $\alpha=30^\circ$ relative to the tank. The camera has a viewing range of $[-4,4]^\circ$. Then $\phi_x$ varies from $-4^\circ$ to $4^\circ$.

@@

Since there is also an uncertainty in the angle, we want $n$ to vary with $\theta_x$ as little as possible. Figures \ref{fig:dndt} show this derivative. The magnitude of this derivative $\left|\partial n/\partial \theta_x\right|$ first decreases with increasing $\theta_x$, reaches a minimum, then increases again. The worst possible angles to measure at are $0^\circ$ and $90^\circ$. The minimum is reached at $31^\circ$. Any angle between $15^\circ$ and $55^\circ$ yields approximately the same sensitivity of $n$.

@@

Concluding, Figures \ref{fig:dndx} and \ref{fig:dndt} indicate that the worst angles to measure are close to zero. Angles between $15^\circ$ and $55^\circ$ result in small errors in $n$.

\section{Forward Model}
\label{sec:formod}
In this section we build a forward model, relating an index of refraction field $n$, position on the ccd-sensor $\underline{x}$ and parameters $\underline{\alpha}$ to the coordinates $\underline{X}$ where the light rays hits the screen. We seek a model $\underline{X}(n, \underline{\alpha}, \underline{x})$.

In Subsection \ref{subsec:rayequ} we derive the equations governing 3-dimensional ray optics in an inhomogeneous medium. In Subsection \ref{subsec:RayPath} give a mathematical description of all the objects our light rays encounter. In Subsection \ref{subsec:dis} we construct the full ray path from the camera through the tank to the screen

@@

\subsection{Mathematical description of Geometrical Objects}
\label{subsec:RayPath}
To give our method flexibility and not making assumption on the relative configuration in our experimental setup, we model the light rays as traversing in a three-dimensional space and allow the different elements in the experimental setup to have any alignments.

@@

\subsubsection{Snell's Law}
To find the angle of incidence, $\theta_I$, between the incoming light ray and the plane, we compute
\begin{equation}
	\cos \theta_I = \underline{\hat{n}} \cdot \underline{I}.
\end{equation}
We apply Snell's law (\ref{eq:snellslaw}) to find the angle of refraction, $\theta_T$. The index of refraction of the medium through which the incoming ray travels is $n_I$, the index of refraction of the medium through wich the outgoing ray travels is $n_T$. To find the direction cosines of the outgoing ray, $\underline{T}$, we calculate 
\begin{align}
	S = \sign \cos \theta_I, &\qquad
	\sin \theta_T = \frac{n_I}{n_T} \sqrt{1-\cos^2 \theta_I}, \\
	\cos \theta_T = \sqrt{1-\left(\frac{n_I}{n_T}\right)^2(1-\cos^2 \theta_I)} &\qquad
	\underline{T} = \frac{n_I}{n_T} \underline{I} + \left(\frac{n_I}{n_T} \cos \theta_I - S \cos\theta_T\right)\underline{\hat{n}},	
\end{align}
where $S$, the sign of $\cos\theta_I$ ensures we stay in the right quadrant.

\subsection{Displacements}
\label{subsec:dis}
Previous subsections combine to give the forward model: $\underline{X}(n,\underline{\alpha}, \underline{x})$. Given an index of refraction field $n$, parameters $\underline{\alpha}$ and coordinates $\underline{x} = (x,y)$, we can calculate the location on the screen where the light rays end up. $\underline{\alpha}$ are the parameters that define the planes $a$, $b$, $c$ and the distance to the screen $L_m$. We use $L_m$ instead of $d$ since we want our parameters to be (linearly) independent. Also, $L_m$ has a clear geometric meaning.
